Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-18 writeup/gz_hubble_data.tex
consistent 'recolored' label in S2

\subsection{Galaxy selection}

In the ACS-GC \citep{gri12}, individual galaxies were identified using a
combination of \sextractor{} \citep{ber96} and the \sersic-profile fitting
framework \galapagos{} \citep{bar12}. GZH included all galaxies with $m<23.5$,
where $m$ is in the \Iband, \zband, or \iband{} for the AEGIS + COSMOS, GEMS +
GOODS-S, and GOODS-N surveys, respectively. This yielded a total of
113,166~images (Table~\ref{tbl:gzh_numbers}).

Single-epoch images from SDSS Stripe~82 were selected using the same criteria
from \citet{wil13}, which required limits of \texttt{petroR90\_r}$ >
3$\arcsec~(where \texttt{petroR90\_r} is the radius containing 90\% of the
$r^\prime$ Petrosian flux) and a magnitude brighter than $m_r < 17.77$. The
images used were three-color composites using the $g^\prime$, $r^\prime$, and
$i^\prime$ filters \citep{nie04}. 21,522 galaxies in SDSS met these criteria.
Co-added images from Stripe~82 were selected from the union of galaxies with
co-added magnitudes brighter than $17.77$~mag, and the galaxies detected in the
single-depth images and matched to a co-add source. This resulted in a total
set of 30,339 images. Of the images in the co-added sample, 5144 (17~percent)
were dimmer than the initial magnitude cut of 17.77.

\subsection{Image creation}

@@

In some cases, we found that attempting to emphasize faint features in the
images resulted in the sky noise taking the appearance of brightly colored
speckles. This impaired the aesthetics of the images and might have been a
distraction to visual classification.  To counteract this, a soft-edged object
mask was applied to the color images and a desaturation operation performed.
This masked procedure was effective in preserving the color balance for
galaxies, retaining the visibility of faint features, but reduced the color
contrast in the sky noise and greatly improved the appearance of the images.
This solution was applied to the coadded Stripe~82, COSMOS and 5-epoch GOODS
images.

\subsection{Images with simulated nuclear point sources}

GZH also included a set of images designed to measure the effect of active
galactic nuclei (AGN) on morphological classifications. Since galactic nuclei
can have bright, unresolved optical emission, AGN have the potential to mimic
or distort the identification of a bulge component. The presence of an AGN was
simulated by modeling the point spread function (PSF) of the telescope and then
inserting a bright source near the center of a real galaxy. For each image, the
simulated AGN was assigned one of three colors -- either blue, red, or flat
(white) as seen in the color images -- and a range of brightnesses such that
$L_\mathrm{ratio} \equiv L_\mathrm{galaxy}/L_\mathrm{AGN}$ is in
$(0.2,1.0,2.0,5.0,10.0,50.0)$. This range was selected to run from AGN which
were all but undetectable through to those which dominate the galaxy, providing
coverage of all plausible effects. Combining these parameters generated
15~images with different simulated AGN for each host.

@@

The previous versions of Galaxy Zoo morphology classifications
\citep{lin08,wil13} were based on observations of galaxies in the Sloan Digital
Sky Survey (SDSS) which have a median redshift of $z<0.2$. In these cases, it
was assumed that there was no cosmological evolution of the morphologies of
galaxies and therefore any observed changes in the distribution of galaxies
with different consensus morphologies was due to the effects of redshift on the
image quality (\ie, the reduction in physical resolution, surface brightness
dimming, etc). For both previous releases of Galaxy~Zoo morphologies, a correction for
redshift-dependent bias was applied based on matching the classification
fractions at the highest redshifts with those at the lowest redshift.
\citet{bam09} and \citet{wil13} provide complete descriptions of the process
for GZ1 and GZ2, respectively.

@@

In order to test and correct for the effects of redshift, GZH includes a set of
calibration images. These images consist of the same galaxy as it would appear
over a variety of redshifts. The input images are from the SDSS
\citep{yor00,str02} and are processed using the \ferengi{} code \citep{bar08a}
to match the observational properties of the \hst{} surveys out to $z=1$. These
images were classified in the GZH interface using the same classification
scheme as the original \hst{} images.

\subsection{Generating images of artificially-redshifted galaxies}

The GZH classifications include 288 unique galaxies originally generated from
SDSS imaging and processed with the \ferengi{} code. The selection spanned a
variety of galaxy morphologies (as selected by GZ2 classifications) and
$r^\prime$-band surface brightnesses, and also spanned the redshift range of
SDSS targets that maximuze the coverage of parameter space.

%The selection criteria for the different morphological categories is summarized in Table \ref{tbl:morphologies}. 
The surface brightness selection ($N_\mu = 3$) for the \ferengi{} galaxies was
(1) low: $\mu > 21.5$~\magarc;  (2) mid: $20.5 < \mu < 21.5$~\magarc; and (3)
high: $\mu < 20.5$~\magarc. For each of the four ``target redshifts''
($z_\mathrm{sim} = 0.3, 0.5, 0.8$ and $1.0$), the images were redshifted in
$\Delta z = 0.1$ bins up to $z_\mathrm{sim}=1.0$.

@@

>>>>

\subsection{The Sloan Digital Sky Survey}\label{ssec:sdss}

The project includes images from the Sloan Digital Sky Survey
\citep[SDSS;][]{yor00,str02}, in particular those from Data~Release~7
\citep{aba09}. A small number of images were used in the creation of simulated
\hst{} images, described in Sections \ref{ssec:ferengi} and \ref{sec:debiasing}
below; the majority of SDSS images used in GZH were from Stripe~82. The
Stripe~82 samples in GZH include single-epoch and co-added images, intended to
facilitate comparison of morphological classification with image depth.

%The creation of 51,861 images from these sources is described in Section
%\ref{ssec:galselect} below.
Single-epoch images from SDSS Stripe~82 were selected using the same criteria
from \citet{wil13}, which required limits of \texttt{petroR90\_r}$ >
3$\arcsec~(where \texttt{petroR90\_r} is the radius containing 90\% of the
$r^\prime$ Petrosian flux) and a magnitude brighter than $m_r < 17.77$. The
images used were three-color composites using the $g^\prime$, $r^\prime$, and
$i^\prime$ filters \citep{nie04}. 21,522 galaxies in SDSS met these criteria.
Co-added images from Stripe~82 were selected from the union of galaxies with
co-added magnitudes brighter than $17.77$~mag, and the galaxies detected in the
single-depth images and matched to a co-add source. This resulted in a total
set of 30,339 images. Of the images in the co-added sample, 5144 (17~percent)
were dimmer than the initial magnitude cut of 17.77.

\subsection{Simulated \hst{} images}\label{ssec:simulatedimages}

To facilitate correction of classifications in the presence of known biasing
effects, GZH includes two different samples of simulated \hst{} images: real
\hst{} galaxy images (Section \ref{ssec:legacy_surveys}) with nuclear emission
added to simulate AGN host galaxies, and lower-redshift galaxy images from SDSS
(Section \ref{ssec:sdss}), artificially redshifted from $0.3 \leq z \leq 1$.
Each is described separately below.

\subsubsection{Images with simulated nuclear point sources}\label{ssec:sim_agn}

As active galactic nuclei often have bright, unresolved optical emission, AGN
have the potential to mimic or distort the identification of a bulge component.
GZH thus includes a set of images designed to measure the effect of AGN on
morphological classifications.  The presence of an AGN was simulated by
modeling the point spread function (PSF) of the telescope and then inserting a
bright source near the center of a real galaxy. For each image, the simulated
AGN was assigned one of three colors -- either blue, red, or flat (white) as
seen in the color images -- and a range of brightnesses such that
$L_\mathrm{ratio} \equiv L_\mathrm{galaxy}/L_\mathrm{AGN}$ is in
$(0.2,1.0,2.0,5.0,10.0,50.0)$. Combining these parameters generated 15~images
with different simulated AGN for each host, in addition to the original galaxy
image.

\subsubsection{Generating images of artificially-redshifted galaxies}\label{ssec:ferengi}

The dimming and resolution effects of redshift can significantly affect galaxy
classifications derived from any method. To facilitate corrections due to
redshift, we include a sample of 288 galaxies with at redshifts such that
available SDSS imaging can be transformed to simulate \hst{} imaging out to $z
= 1$. The selection spanned a variety of galaxy morphologies (as selected by
GZ2 classifications) and $r^\prime$-band surface brightnesses, and also spanned
the redshift range of SDSS targets (in $N_z = 4$ bins) in order to optimize for
different target minimum redshifts in \hst{} imaging. We redshifted the
galaxies and produced simulated \hst{} imaging using the \ferengi{} code
\citep{bar08a}.

%The selection criteria for the different morphological categories is
%summarized in Table \ref{tbl:morphologies}. 
The surface brightness selection ($N_\mu = 3$) for the \ferengi{} galaxies was
(1) low: $\mu > 21.5$~\magarc;  (2) mid: $20.5 < \mu < 21.5$~\magarc; and (3)
high: $\mu < 20.5$~\magarc. For each of the four ``target redshifts''
($z_\mathrm{sim} = 0.3, 0.5, 0.8$ and $1.0$), the images were redshifted in
$\Delta z = 0.1$ bins up to $z_\mathrm{sim}=1.0$.

In addition to the physical parameters of the input images, the \ferengi{}
output depends on assumptions of the global galaxy evolution model. This
evolution is a crude mechanism that mimics the brightness increase of galaxies
with increasing redshift (out to at least $z\sim1-2$). The effect on the
redshifted images is simply an empirical addition to the magnitude of a galaxy
of the form $M' = e\times z + M$, where $M'$ is the corrected magnitude, and
$e$ is the evolutionary correction in magnitudes (i.e., $e=-1$ essentially
brightens the galaxy by 1~magnitude by $z=1$). \ferengi{} was run on the images
for values of $e$ starting from $e=0$ and decreasing to $e=-3.5$ in increments
of $\Delta e = 0.5$. Figure~\ref{fig:exampleFERENGI} shows several examples
of the effects of ``losing'' spiral/disc features with increasing redshift
for two galaxies with $e=0$.

The final number of \ferengi{} images produced for each galaxy is ultimately a
function of galaxy's redshift (since the new images cannot be resampled at
better angular resolution than the original SDSS data), as well as the number
of $e$ values selected. Table~\ref{tbl:ferengivalues} summarizes the redshifted
images produced for GZH. The use of these images to correct for redshift-dependent
classification bias is described in Section \ref{sec:debiasing}.

\subsection{Image creation}\label{ssec:images}

@@

In some cases, we found that attempting to emphasize faint features in the
images resulted in the sky noise taking the appearance of brightly colored
speckles. This impaired the aesthetics of the images and might have been a
distraction to visual classification.  To counteract this, a soft-edged object
mask was applied to the color images and a desaturation operation performed.
This masked procedure was effective in preserving the color balance for
galaxies, retaining the visibility of faint features while reducing the color
contrast in the sky noise and greatly improving the appearance of the images.
This solution was applied to the coadded Stripe~82, COSMOS and 5-epoch GOODS
images.

In addition to the main \hst{} legacy imaging sample with colored images
described as above, GZH also includes samples of \hst{} galaxy images with
different color prescriptions. In particular, there are 2 sets of 3,927 images
each, drawn from the COSMOS sample; the first has a dramatically reduced color
saturation, and the second has reversed the colors so that the blue and red
filters have exchanged places in the RGB image. The `faded' galaxy set
facilitates measuring of possible variations in classification due to the
presence or absence of color features. The `recolored' set enables an
alternative test of potential color biases in classifications.
% I guess this means we need to say somewhere later that we don't see an effect
% (we don't, right?).  Would also be good to cite here where we have done
% similar things in the past...

The simulated AGN host galaxy color images (described in Section \ref{ssec:sim_agn})
were created using the same prescription as for the GOODS 2-epoch imaging. The
artificially redshifted color images (Section \ref{ssec:ferengi}) were created using the
same prescription as for the AEGIS images.
% Edmond should verify the AEGIS bit above.

\subsection{Galaxy sample labels}\label{ssec:sample_labels}

The full GZH sample is composed of 7~different galaxy sub-samples. Throughout
this paper and in the published catalogs they are referred to with the
following labels:

%\ref{tbl:catalog_hst}
\item \texttt{main}: \hst{} imaging with RGB colors making use of all available
filters with typical saturation and in correct order, as described in
Section~\ref{ssec:images}. This sample includes AEGIS, COSMOS, GEMS, and
full-depth GOODS (North and South) images. (118,425 galaxies)

%\ref{tbl:catalog_faded}
\item \texttt{faded}: a subset of COSMOS images with very low color saturation.
(3,927 galaxies)

%\ref{tbl:catalog_recolored}
\item \texttt{recolored}: a subset of COSMOS images with red and blue channels
reversed. Note: this sub-sample uses the same galaxies as the \texttt{faded}
sub-sample. (3,927 galaxies)

%\ref{tbl:catalog_goods_shallow}
\item \texttt{goods-shallow}: Images from GOODS-North and GOODS-South, with
color from 2 filters and imaged at 2-epoch depth. (6,144 galaxies)

%\ref{tbl:stripe82_single}
\item \texttt{stripe-82-single}: Single-epoch images from SDSS Stripe~82.
(21,522 galaxies)

%\ref{tbl:stripe82_coadd}
\item \texttt{stripe-82-coadd}: Co-added images from SDSS Stripe~82. Note: this
sub-sample includes all galaxies in the \texttt{stripe-82-single} sub-sample,
with additional sources detected in the deeper imaging. (30,339 galaxies)

%ferengi <-- no table?
\item \texttt{redshifted}: Simulated \hst{} images constructed using SDSS
images and artificially redshifted to between $0.3 \leq z \leq 1$. (288
original galaxies; 3,950 redshifted galaxies)

%\ref{tbl:simulated_agn}
\item \texttt{simulated-agn}: Simulated AGN host galaxies constructed using
\hst{} images and PSFs. (96 original galaxies; 2,961 simulated AGN host
galaxies)

@@

The previous versions of Galaxy Zoo morphology classifications
\citep{lin11,wil13} were based on observations of galaxies in the Sloan Digital
Sky Survey (SDSS) which have a median redshift of $z<0.2$. In these cases, it
was assumed that there was no cosmological evolution of the morphologies of
galaxies and therefore any observed changes in the distribution of galaxies
with different consensus morphologies was due to the effects of redshift on the
image quality (\ie, the reduction in physical resolution, surface brightness
dimming, etc). For both previous releases of Galaxy~Zoo morphologies, a correction for
redshift-dependent bias was applied based on matching the classification
fractions at the highest redshifts with those at the lowest redshift.
\citet{bam09} and \citet{wil13} provide complete descriptions of the process
for GZ1 and GZ2, respectively.

@@

In order to test and correct for the effects of redshift, GZH includes a set of
calibration images. These images consist of the same galaxy as it would appear
over a variety of redshifts. The input images are from the SDSS
\citep{yor00,str02} and are processed to match the observational properties of 
the \hst{} surveys out to $z=1$. These
images were classified in the GZH interface using the same classification
scheme as the original \hst{} images. Further details on the creation of these images
is given in Section \ref{ssec:ferengi}.

The GZH classifications include 288 unique galaxies originally generated from
SDSS imaging and processed with the \ferengi{} code. The selection spanned a
variety of galaxy morphologies (as selected by GZ2 classifications) and
$r^\prime$-band surface brightnesses, and also spanned the redshift range of
SDSS targets that maximize the coverage of parameter space.

%The selection criteria for the different morphological categories is
%summarized in Table \ref{tbl:morphologies}. 
The surface brightness selection ($N_\mu = 3$) for the \ferengi{} galaxies was
(1) low: $\mu > 21.5$~\magarc;  (2) mid: $20.5 < \mu < 21.5$~\magarc; and (3)
high: $\mu < 20.5$~\magarc. For each of the four ``target redshifts''
($z_\mathrm{sim} = 0.3, 0.5, 0.8$ and $1.0$), the images were redshifted in
$\Delta z = 0.1$ bins up to $z_\mathrm{sim}=1.0$.

@@

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/example_ferengi.pdf}
\caption{Examples of two galaxies which have been run through the \ferengi{}
code to produce simulated \hst{} images. The measured value of \ffeatures{}
from GZH for the images in each panel are (1) Top row: \ffeatures~= (0.900,
0.625, 0.350, 0.350, 0.225) and (2) Bottom row: \ffeatures~= (1.000, 0.875,
0.875, 0.625, 0.375). \label{fig:exampleFERENGI}}
\end{figure*}
>>>>>>> master
